On Sample Size Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2024-01-09 Manuscript/manuscript.tex
touch up

For estimating the mean parameter~$\mu$, \textbf{Figure~\ref{fig:norm1}} suggests 
that all methods eventually approach correct coverage of $\mu$ as sample size 
increases. Student's $t$ CIs appear to need the smallest sample size to achieve 
correct coverage, except for samples with strong negative dependence, in which 
case, they actually over-cover $\mu$ for smaller sample sizes. For instance, for 
a sample with $n = 100$ and $\phi = -0.4$, the lower bound for a Students $t$ 
CI's coverage of $\mu$ is greater than 95\%, whereas the coverage intervals for 
other methods contain 95\%. The standard normal, percentile, BC, and BCA, and 
recentered percentile CIs require similar sample sizes to recover $\mu$ at the 
nominal level for all combinations of $n$ and $\phi$. All methods seem to 
require a smaller sample to recover $\mu$ at the nominal rate when dealing with 
negative dependence versus positive dependence. For example, BC CIs recover 
$\mu$ for $n \geq 100$ when $\phi = -0.2$, but they only recover $\mu$ for 
$n \geq 800$ when $\phi = 0.2$. In addition, as a negative dependence gets 
stronger, holding everything else equal, coverage increases, which lead to the 
Student $t$ CI's aforementioned over-coverage. As a positive dependence gets 
stronger, holding everything else equal, coverage decreases, and a larger sample 
is necessary to recover~$\mu$. A possible explanation for this is that if a 
stationary series has a positive 
autocorrelation, the effective sample
size is decreased, whereas if a series has a negative autocorrelation, the
effective sample size is increased. \citep{geyer2011introduction} Additionally, 
this seems to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter.

@@

Block bootstrap is a useful method for estimating parameters of a time series,
from simple parameters like the mean to more complicated temporal dependence
factors. We know theoretically that the block bootstrap procedure will cover a 
parameter of a time series at the nominal level given an infinitely large 
sample,\citep{calhoun2018} so the goal for this study was to find the smallest 
finite sample length~$n$
of a time series in order for the block bootstrap procedure to recover its 
associated parameters at an acceptable rate. Our analysis relies on the 
assumption that there is a size $n$ large enough for the method to work: that 
is, the method's performance improves as $n$ increases. Out of the six types of 
intervals used in this study, this assumption was found to hold true with 
respect to estimating $\phi$ only for standard normal, Student's $t$, and recentered 
percentile CIs, whereas percentile, BC, and BCA intervals exhibited coverage 
deterioration as $n$ increased. The percentile CI's coverage deterioration can 
be attributed to bias that is not corrected as $n$ increases. Specifically, as 
$n$ increases, the width of the CI decreases, but because the percentile CI 
underestimates $\phi$, the coverage decreases. The BC CI seems to correct the 
bias, but the width of the CI seems to be too short. The acceleration factor 
of the BCA CI seems to fail, as the width of the CI seems to be too
short.

>>>>

For estimating the mean parameter~$\mu$, \textbf{Figure~\ref{fig:norm1}} suggests 
that all methods eventually approach correct coverage of $\mu$ as sample size 
increases. Student's $t$ CIs appear to need the smallest sample size to achieve 
correct coverage, except for samples with strong negative dependence, in which 
case, they actually over-cover $\mu$ for smaller sample sizes. For instance, for 
a sample with $n = 100$ and $\phi = -0.4$, the lower bound for a Students $t$ 
CI's coverage of $\mu$ is greater than 95\%, whereas the coverage intervals for 
other methods contain 95\%. The standard normal, percentile, BC, and BCA, and 
recentered percentile CIs require similar sample sizes to recover $\mu$ at the 
nominal level for all combinations of $n$ and $\phi$. All methods seem to 
require a smaller sample to recover $\mu$ at the nominal rate when dealing with 
negative dependence versus positive dependence. For example, BC CIs recover 
$\mu$ for $n \geq 100$ when $\phi = -0.2$, but they only recover $\mu$ for 
$n \geq 800$ when $\phi = 0.2$. In addition, as a negative dependence gets 
stronger, holding everything else equal, coverage increases, which lead to the 
Student $t$ CI's aforementioned over-coverage. As a positive dependence gets 
stronger, holding everything else equal, coverage decreases, and a larger sample 
is necessary to recover~$\mu$. A possible explanation for this is that if a 
stationary series has a positive 
autocorrelation, the effective sample
size is decreased, whereas if a series has a negative autocorrelation, the
effective sample size is increased.\citep{geyer2011introduction}
Additionally,  this seems to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter.

@@

Block bootstrap is a useful method for estimating parameters of a time series,
from simple parameters like the mean to more complicated temporal dependence
factors. We know theoretically that the block bootstrap procedure will cover a 
parameter of a time series at the nominal level given an infinitely large
\jy{Also cite Calhoun in the introduction}
sample,\citep{calhoun2018} so the goal for this study was to find the smallest 
finite sample length~$n$
of a time series in order for the block bootstrap procedure to recover its 
associated parameters at an acceptable rate. Our analysis relies on the 
assumption that there is a size $n$ large enough for the method to work: that 
is, the method's performance improves as $n$ increases. Out of the six types of 
intervals used in this study, this assumption was found to hold true with 
respect to estimating $\phi$ only for standard normal, Student's $t$, and recentered 
percentile CIs, whereas percentile, BC, and BCA intervals exhibited coverage 
deterioration as $n$ increased. The percentile CI's coverage deterioration can 
be attributed to bias that is not corrected as $n$ increases. Specifically, as 
$n$ increases, the width of the CI decreases, but because the percentile CI 
underestimates $\phi$, the coverage decreases. The BC CI seems to correct the 
bias, but the width of the CI seems to be too short. The acceleration factor 
of the BCA CI seems to fail, as the width of the CI seems to be too
short.
